\myTitle

### 2014-08-13 tex/main-body-text.tex
Made final edits to manuscript and readme.

\subsection*{{\sc flock}}
Here we translate the recipe given by \citeauthor{Duc&Tur2009} for the 
{\sc flock} algorithm into a specification in terms of the variables
defined in the previous section.  We provide the description for a case in which
it is assumed there are $K$ clusters.

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{images/Figures-Pat/Flock-Fig1.pdf}   % Use this if we want it grayscale -> Flock-Fig1-GreyScale.pdf  
    \caption{Plots of $q_i$ values from {\sc structure} and the normalized likelihood values 
from {\sc flock} made with the program {\sc distruct}. Each horizontal plot represents an individual run of the program {\sc structure}, or the `best run' from a series of runs in {\sc flock}. Within each plot 
$q_i$ values from {\sc structure} and the normalized likelihood values 
from {\sc flock}  are represented by a 
gray-scale vertical bar. The numbers in parentheses on the right indicate how many {\sc structure} runs (out of 6, as indicated) reached the same visually indistinguishable solution, or how many groups of {\sc flock} runs (out of 6) had the same `best run.'}
    \label{fig:qplots}
\end{center}
\end{figure*}

@@

Though we observed similar results
between the two programs at low values of $K$, {\sc flock}'s `best-run' solutions became somewhat inconsistent at higher values of $K$.
This is likely a consequence of the rapidly increasing size of the search space over possible 
partitions as $K$ and $n$, the number of individuals, increase. 
The likelihood surface defined on partitions of this space will often
be multimodal---possessing numerous local peaks and troughs.
The peak that  {\sc flock}'s deterministic algorithm finds is entirely determined by the initial allocation of 
individuals, as it is a deterministic algorithm.
If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it less likely that many (if any at all) will converge to the same local mode. This coupled with the fact that there is 
ambiguity in what defines a `best run' when all plateaus are of length zero complicates the interpretation
of results from {\sc flock} if no non-zero plateaus are encountered. The problem of getting 
caught in local modes is not unique to {\sc flock}---{\sc structure} also can converge to 
different solutions in complex spaces; however our results suggest this is a bigger problem for {\sc flock}.

One of the advantages of {\sc flock} over other methods for genetic clustering
is reported to be the short processing time for each run of {\sc flock}.
While \citet{Duc&Tur2012} observed faster processing times (relative to other methods) in their 
comparisons, we did not observe this advantage in our analysis. This appears to result from two issues. First,
our example data set was an order of magnitude larger than the largest data set analyzed in \citet{Duc&Tur2012}.  
Second, in the comparison by \citet{Duc&Tur2012}, 
{\sc flock} was set to perform 50 runs and 20 iterations for each $K$ starting at 2 and ending when 
one of the stopping conditions was met, and computing times for that analysis were compared to ten iterations of {\sc structure} using 
a 50,000 sweep burn-in followed 
by a 200,000 sweep sample from the posterior for values of $K$ from 1 to three more than the true 
value of $K$. 
To us it seemed reasonable to compare, for each $K$, the running times of one iteration of {\sc structure} (50,000 
burn in and 200,000 sample collection sweeps)
to one iteration of {\sc flock} (50 different starting allocations that are reallocated 20 times) 
as both were used to provide one of 6 total estimates (per value of $K$) of the partition of individuals into 
clusters. Though
it is difficult to reliably assess differences in program run times when there is no clear consensus on what 
constitutes a reasonable comparison with respect to run length and number of runs, our results indicate that {\sc flock} did not carry a distinct run-time advantage over {\sc structure}.

@@

Often, newly-developed algorithms come to be understood within more general contexts and 
frameworks, and this, in turn, can help guide improvement of those algorithms.  For example, in population
genetics, the ``Markov recursion'' algorithm of \citet{Gri&Tav1994-AI} was identified by \citet{Felsensteinetal1999} 
as a special case of importance sampling \citep{Ham&Han1964}, a perspective which allowed 
\citet{Ste&Don2000} to improve upon the algorithm dramatically.  Here we have identified the {\sc flock}
algorithm as a limiting case of simulating annealing upon {\sc structure}'s no-admixture model with uncorrelated
allele frequencies. We hope this will
help users to understand and interpret {\sc flock}'s behavior and will provide insights that may be useful 
to {\sc flock}'s authors if they further develop their algorithm.

>>>>

\subsection*{{\sc flock}}
Here we translate the procedure given by \citeauthor{Duc&Tur2009} for the 
{\sc flock} algorithm into a specification in terms of the variables
defined in the previous section.  We provide the description for a case in which
it is assumed there are $K$ clusters.

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{images/Figures-Pat/Flock-Fig1.pdf}   % Use this if we want it grayscale -> Flock-Fig1-GreyScale.pdf  
    \caption{Plots of $q_i$ values from {\sc structure} and the normalized likelihood values 
from {\sc flock} made with the program {\sc distruct}. Each horizontal plot represents an individual run of the program {\sc structure}, or the `best run' from a series of runs in {\sc flock}. Within each plot 
$q_i$ values from {\sc structure} and the normalized likelihood values 
from {\sc flock}  are represented by a 
%gray-scale
vertical bar. The numbers in parentheses on the right indicate how many {\sc structure} runs (out of 6, as indicated) reached the same visually indistinguishable solution, or how many groups of {\sc flock} runs (out of 6) had the same visually indistinguishable `best run.'}
    \label{fig:qplots}
\end{center}
\end{figure*}

@@

Though we observed similar results
between the two programs at low values of $K$, {\sc flock}'s `best-run' solutions became somewhat inconsistent at higher values of $K$.
This observation was consistent with \citet{Duc&Tur2012}'s comparison of the 
two programs using simulated data (Fig. 4 in \citet{Duc&Tur2012})
for $K=4 \& 8, L=10$. 
This is likely a consequence of the rapidly increasing size of the search space over possible 
partitions as $K$ and $n$, the number of individuals, increase. 
The likelihood surface defined on partitions of this space will often
be multimodal---possessing numerous local peaks and troughs.
The peak that  {\sc flock}'s deterministic algorithm finds is entirely determined by the initial allocation of 
individuals, as it is a deterministic algorithm.
If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it less likely that many (if any at all) will converge to the same local mode. 
This coupled with the fact that there is 
ambiguity in what defines a `best run' when all plateaus are of length zero complicates the interpretation
of results from {\sc flock} if no non-zero plateaus are encountered. The problem of getting 
caught in local modes is not unique to {\sc flock}---{\sc structure} also can converge to 
different solutions in complex spaces; however our results suggest this is a bigger problem for {\sc flock}.
By increasing the number of loci used in their simulations \citet{Duc&Tur2012} showed, surprisingly, that 
{\sc flock} did converge on $K=8$, but held no advantage in inference of $K$ over {\sc structure}. We suspect that this 
resulted from the confined search space defined
by the small number of individuals (n=240) used in the analysis. 
%They simulated data using 500 individuals per population but only used 30 of them in the analysis.
% if you have more loci than individuals per population with high Fst between groups it does great!

One of the advantages of {\sc flock} over other methods for genetic clustering
is reported to be the short processing time for each run of {\sc flock}.
While \citet{Duc&Tur2012} observed faster processing times (relative to other methods) in their 
comparisons, we did not observe this advantage in our analysis. This appears to result from two issues. First,
our example data set was an order of magnitude larger than the largest data set analyzed in \citet{Duc&Tur2012}.  
Second, in the comparison by \citet{Duc&Tur2012}, 
{\sc flock} was set to perform 50 runs and 20 iterations for each $K$ starting at 2 and ending when 
one of the stopping conditions was met, and computing times for that analysis were compared to ten iterations of 
the admixture and correlated allele frequency model in {\sc structure} using 
a 50,000 sweep burn-in followed 
by a 200,000 sweep sample from the posterior for values of $K$ from 1 to three more than the true 
value of $K$. 
To us it seemed reasonable to compare, for each $K$, the running times of one iteration of {\sc structure} (50,000 
burn in and 200,000 sample collection sweeps)
to one iteration of {\sc flock} (50 different starting allocations that are reallocated 20 times) 
as both were used to provide one of 6 total estimates (per value of $K$) of the partition of individuals into 
clusters. Though
it is difficult to reliably assess differences in program run times when there is no clear consensus on what 
constitutes a reasonable comparison with respect to run length and number of runs, our results indicate that {\sc flock} did not carry a distinct run-time advantage over {\sc structure}.

@@

Often, newly-developed algorithms come to be understood within more general contexts and 
frameworks, and this, in turn, can help guide improvement of those algorithms.  For example, in population
genetics, the ``Markov recursion'' algorithm of \citet{Gri&Tav1994-AI} was identified by \citet{Felsensteinetal1999} 
as a special case of importance sampling \citep{Ham&Han1964}, a perspective which allowed 
\citet{Ste&Don2000} to improve upon the algorithm dramatically.  Here we have identified the {\sc flock}
algorithm as a limiting case of simulating annealing upon {\sc structure}'s no-admixture model with uncorrelated
allele frequencies. We hope this will
help users to understand and interpret {\sc flock}'s behavior and will provide insights that may be useful 
in the potential development of the {\sc flock} algorithm.
